### 2024-06-02 2024-5-13/sections/heuristic.tex
fix the syntax error in heuristics

%The naive approach for nested counting in the paper \cite{Denghang2023} is to unwind all inner counting operators. This approach may generate a large number of states, which is linear to the product of the counting bounds in regex and exponential to the length of the regex. For example, naive unwinding regex $(a^{\{1,1000\}})^{\{1,2\}}$ generate a CEFA containing 1000 states. But if we only unwind the outer counting operator of the regex, not inner, we obtain a CEFA containing only 2 states. The side effect is that we need two registers to store the information of conuting in the inner regex (unwinding of outer counting operator repeat the inner regex, which also repeat the registers of the inner regex).We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to unwind the counting operator with less bound and less side effect firstly. We present the heuristic in the following algorithm.

@@

For a regex $\regex$, we can approximate the number of states $size(\regex)$ generated by the following formula: 
\begin{itemize}
  \item If $\regex = \emptyset$ or $\regex = \epsilon$, then $size(\regex) = 0$. 
  \item If $\regex = \regex_1 \cdot \regex_2$ or $\regex = \regex_1 + \regex_2$, then $size(\regex) = size(\regex_1) + size(\regex_2)$.
  \item If $\regex = \regex_1^*$, then $size(\regex) = size(\regex_1)$.
  \item If $\regex = \regex_1 \cap \regex_2$, then $size(\regex) = size(\regex_1)*size(\regex_2)$.
  \item If $\regex = \overline{\regex_1}$, then $size(\regex) = 2^{size(\regex_1)} + 1$.
  \item $\regex = \regex_1 \backslash \regex_2$ can be transfered to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
  \item For $\regex = \regex_1^{\{n,m\}}$ or $\regex = \regex_1^{\{m,\infty\}}$, $size(\regex)$ is equal to $m*size(\regex_1)$ if it is unwinded, and $size(\regex_1)$ otherwise.
\end{itemize}

@@

We unwind the counting operator with less bound first. The reason is that the number of states generated by the counting operator is linear to its bound. If we unwind the counting operator with less bound first, the number of states generated by the unwinding process is less. However, the hardness of solving the $\cefadec$ problem is not only relevant to the number of states, but also the number of registers. For a regex $\regex$, we can approximate the number of registers $reg(\regex)$ generated by the following formula:
\begin{itemize}
  \item If $\regex = \emptyset$, $\regex = \epsilon$, $\regex = \regex_1^*$ or $\regex = \overline{\regex_1}$, then $reg(\regex) = 0$.
  \item If $\regex = \regex_1\cdot \regex_2$ or $\regex = \regex_1 \cap \regex_2$, then $reg(\regex) = reg(\regex_1) + reg(\regex_2)$.
  \item If $\regex = \regex_1 + \regex_2$, then $reg(\regex) = reg(\regex_1) + reg(\regex_2) + 1$.
  \item $\regex = \regex_1 \backslash \regex_2$ can be transfered to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the number of registers can be computed by the formula of $\cap$ and $\bar{}$.
  \item For $\regex = \regex_1^{\{n,m\}}$ or $\regex = \regex^{\{m, \infty\}}$, $reg(\regex)$ is equal to $m*reg(\regex_1)$ if it is unwinded, and $1$ otherwise.
\end{itemize}

@@

\medskip
\noindent
\emph{Step 3. (Unwind the counting operator with less bound and less registers first).}

In the above two steps, we compute the number of states and registers generated by each counting operator. In some cases, we can unwind the counting operator with less bound and less registers directly. But in most cases, there are multiple counting operators where one counting operator has less bound and another counting operator has less registers. To consider the number of states and registers at the same time, we can compute the score of each counting operator by the following formula:
\begin{itemize}
  \item $score(\regex) = size(\regex) * (reg(\regex)+1)$.
\end{itemize} 
The reason why we use product to combine the two factors is that the hardness of the $\cefadec$ problem is highly relevant to the number of variables in the final existential LIA formula, which is linear to the product of the number of states and the number of registers. Use $reg(\regex) + 1$ not $reg(\regex)$ is because the registers number may be $0$.We unwind the counting operator with the smallest score first.

\begin{example}
  Consider the regex $(a^{\{1,1000\}})^{\{1,2\}}$. The score of $^{\{1,1000\}}$ is $1000*(0+1) = 1000$ and the score of $^{\{1,2\}}$ is $2*(2+1) = 6$. We unwind the outer counting operator first and obtain a CEFA containing only 2 states.  
\end{example}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As mentioned in Section~\ref{sec:algorithm}, if some counting operators are in the scope of a complement operator, it is necessary to unfold these counting operators before the complement operation. The exponential blow-up in the unfolding combined with the blow-up in the complement operation will produce huge number of states.  To alleviate this problem, we propose two approximations of the complement operation, i.e. an under-approximation and an over-approximation, and combine them into a procedure that achieves a nice balance between efficiency and precision.

@@

\paragraph*{Under-approximation}
%
One way of doing the complement operation is to transform $e_1$ into a regular expression that does not contain counting operators. We compute an under-approximation of $e$ by replacing each occurrence of the counting operators in $e_1$, say $(e_2)^{\{m,n\}}$ or $(e_2^{\{n, \infty\}})$, by $e_2^*$. Let us use $e'_1$ denote the resulting regular expression. Note that $e'_1$ does not contain counting operators. Moreover, it is easy to observe that $\Lang(e_1) \subseteq \Lang(e'_1)$. As a result, we have $\Lang(\overline{e'_1}) \subseteq \Lang(\overline{e_1})$. That is, $\overline{e'_1}$ is an under-approximation of $\overline{e_1}$.

%The under-approximation is replacing each counting operator in the subregex of complement with a Kleene star operator. The language of the subregex is a superset of the language of it after under-approximation. In sequential, the complement of the subregex is a subset of origin. Note that this approach only works for non-nested complement, which means there is no other complement operators in the complement. For nested complement,  we naively handle the inner complement operators as paper \cite{Denghang2023} and then under-approximate the outer complement operator.

@@

Without loss of generality, we can assume that there are no transitions out of $F$ in $\aut_{e_1}$. If this is not the case, then a new state $q_f$ can be introduced to transform $\aut_{e_1}$ into a CEFA satisfying this constraint, where $q_f$ is the unique accepting state. 
Then we define $\cB= (R \cup \{r'\}, Q \cup \{q_s\}, \Sigma, \delta', I, Q \cup \{q_s\}, \alpha')$, where
\begin{itemize}
  \item $r'$ is a new register not in $R$, 
  \item $q_s$ is a new state not in $Q$,
  \item $\delta'$ comprises the following transitions, 
  \begin{itemize}
    \item $(q, a, q', (\vec{v}, 0))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \not \in F$,
%
    \item $(q, a, q_s,(\vec{0}, 0))$ such that there do \emph{not} exist $a$-labeled transitions out of $q$, that is, for every $q'$ and $\vec{v}$, we have $(q, a, q', \vec{v}) \not\in \delta$,
%
    \item $(q, a, q', (\vec{v}, 1))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \in F$,
%
    \item $(q_s, a, q_s, (\vec{0}, 0))$ such that $a \in \Sigma$, 
  \end{itemize}
  \item $\alpha'$ is defined as $r' = 1 \rightarrow \neg \alpha$.
\end{itemize}
Intuitively, $q_s$ is a sink state that deals with the situations that there are no runs and $r'$ is a register newly introduced so that the different cases of the non-accepting runs of $\aut_{e_1}$ can be captured by a single accepting condition $\alpha'$.

@@

The under-approximation is replacing each counting operator in the subregex of complement with a Kleene star operator. The language of the subregex is a superset of the language of it after under-approximation. In sequential, the complement of the subregex is a subset of origin. Note that this approach only works for non-nested complement, which means there is no other complement operators in the complement. For nested complement,  we naively handle the inner complement operators as paper \cite{Denghang2023} and then under-approximate the outer complement operator.

@@

$r$ is the register used to check whether the accepted state is also accepted in the original CEFA. $s$ is the self loop state accepting all string that is not accepted by the original CEFA. $f$ is the merged accepted state of original CEFA. If there is a transition $(q, a, q', \vec{v})$ teminates at the accepted state $q'$ in the original CEFA, we add a corresponding transition $(q, a, f, (\vec{v}, 1))$ in the over-approximated complement. From the definition of $\delta'$, we can see that only the transitions goto $f$ increase the value of regesiter $r$. The value of $r$ is $1$ if and only if the accepted state is also accepted in the original CEFA. With above construction, the meaning of the accepting condition $\alpha'$ is obvious: If the string teminates at the accepting state in the original CEFA, then the accepting condition $\alpha$ is false.

@@

   In the definition of CEFA in section \ref{sec:automaton}, a run $q_0 \xrightarrow[\myvec{\overline{v}_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{\overline{v}_n}]{a_n} q_n$ is \emph{accepting} by $\overline{\aut}_{over}$ if $q_n \in (Q\cup\{s,f\})\backslash F$ and $\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ is true, where $\myvec{\overline{v}'} = \sum \limits_{j \in [n]} \myvec{\overline{v}_j}$. If a string is not accepted by $\overline{\aut}_{over}$, then each run of the string is not accepting run, which means the run teminates at state $q\in F$ or the accepting condition $\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ is $false$.

@@

   If each run makes the accepting condition $\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ false, then the negation of the accepting condition $\neg\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ is true. The negation can be rewriten to $(r=1\wedge \alpha)(\myvec{\overline{v}'}/(R\cup\{r\})) = (r=1)(\myvec{\overline{v}'}/(R\cup\{r\}))\wedge \alpha(\myvec{\overline{v}'}/(R\cup\{r\}))$. In our construction, only the transitions teminate at $f$ increase one to the value of $r$, so that the run must end at $f$ to make $r=1$ be true. $\alpha(\myvec{\overline{v}'}/(R\cup\{r\})) = \alpha(\myvec{v'}/R)$ is also need to be true, implying that the run is an accepting run of $\aut$.
\end{proof}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Recall that in Section~\ref{sec:algorithm}, the satisfiability of the original RECL constraint is reduced to the satisfiability of the existential LIA formula
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
Let us use $\theta$ to denote this formula. 
Then we utilize some SMT solver, say Z3, to solve the satisfiability of $\theta$. If Z3 reports ``sat'', then it also returns a model, namely, an assignment $\amodel$ of natural numbers to the free integer variables in $\theta$. Then we utilize $\amodel$ to generate a model $\amodel'$ of the original RECL constraint $\varphi$, i.e. an assignment of strings to the string variables so that $\varphi$ is evaluated to true under $\amodel'$.

@@

It turns out that this model generation process is inefficient, mainly due to the fact that the information about the occurrences of vectors $\myvec{v}$ in $\cB_i$ in the searching process is too coarse. In the sequel, we would like to refine this process by using the LIA formulas to obtain the information about the number of occurrences of transitions, instead of just the occurrences of integer vectors, and utilize this information to guide the search process. 
Specifically, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, we introduce an integer variable $\anivarz_{i, (q, a, q', \myvec{v})}$ to denote the number of occurrences of the transition, and solve the satisfiability of the following formula,  
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right]\ \wedge \\
\ \ \bigwedge \limits_{i \in [n]} \bigwedge \limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} = \sum \limits_{(q, a, q', \myvec{v}) \mbox{ occurs in } \cB_i} \anivarz_{i, (q, a, q', \myvec{v})}.
\end{array}
\]
If an SMT solver reports that the aforementioned formula is satisfiable, then the solver can also returns a model, which assigns to each variable $\anivarz_{i, (q, a, q', \myvec{v})}$ a natural number. 
%For each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, let us use $nt_{i, (q, a, q', \myvec{v})}$ to denote the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model.
Then we utilize these values to generate a model (an assignment of strings to string variables) of $\varphi$. 
Let $i \in [n]$. We use a function $\lambda_i$ to record the number of occurrences of the transitions in $\cB_i$. During the search process for $\cB_i$, each time when a transition $(q, a, q', \myvec{v})$ is traversed, $\lambda_i$ is updated by decrementing $\lambda_i((q, a, q', \myvec{v}))$, moreover, $a$ is also recorded. Initially, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, $\lambda_i((q, a, q', \myvec{v}))$ is set to be the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model. When a final state of $\cB_i$ is reached and $\lambda_i((q, a, q', \myvec{v}))$ becomes zero for each transition $(q, a, q', \myvec{v})$ in $\cB_i$, then a desired string is generated for the string variable $x_i$.

Moreover, propose a method to prune the search space further. Let $i \in [n]$. Suppose that during the search process for $\cB_i$, let $q$ be the current state of $\cB_i$ and $\lambda_i$ record the remaining occurrences of transitions of $\cB_i$. We compute the set of transitions of $\cB_i$ that are reachable from the state $q$. Let $\cT_q$ denote this set. If $\{(q, a, q', \myvec{c}) \mid \lambda_i((q, a, q', \myvec{c})) > 0\} \not \subseteq \cT_q$, then we know that it is impossible to continue the search process so that a string accepted by $\cB_i$ will be found and meanwhile the constraint on the number of transitions enforced by $\lambda_i$ is satisfied. Therefore, in this case, the search process for $\cB_i$ in this direction halts and the search process is backtracked.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
In the section \ref{sec:algorithm}, we present an algorithm to solve the $\cefadec$ problem. The algorithm computes existential LIA formulas and then uses SMT solver to solve it. However, this procedure only anwser the question whether the language of the regex is empty or not. If the language is not empty, we need to find a word in the language in practice. In our preivous paper \cite{Denghang2023}, we do not explicitly present the algorithm to find a word in the language because of the space limitation. We have used a naive approach which randomly searching the CEFAs by the depth-first search, based on the regesiters values obtained from solving the existential LIA formula. However, this approach is not efficient when the regexes is non-regesiter-representable. Suppose that there are $n$ registers whose values are $r_1,\cdots, r_n$ respectively in the CEFA, we may visit each transition $O(r_1*r_2\cdots *r_n)$ times during the search. To relieve this problem, we present a heuristic to find a word in the language as follows.

@@

For each CEFA $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$, we already know the values of its registers after solving the existential LIA formula. The values can be seen as a constant vector of length $|R|$. We use $\myvec{R}$ to denote the vector and use $\eta((q, a, q', \myvec{v}))$ to denote the parikh image of $\aut$. Then the transitions times visited by the accepted stirng is obtained from the following formula:
$$ \sum \limits_{(q, a, q', \myvec{v})\in \delta} \eta((q, a, q', \myvec{v}))\myvec{v} = \myvec{R}$$

\medskip
\noindent
\emph{Step 2. (Prune the search tree early by the transitions times during the search process).}

After computing the transitions times visited by the accepted string, we prune the search tree of the CEFA based on the transitions times. The idea is that we only visit the state which is able to reach all transitions need to be visited. Other states are deleted from the CEFA so that we can reduce the search space.
Note that the transitions times is updated during the search process so the search space is reduced dynamically. The presudo code is presented in Algorithm~\ref{alg:findword}.

>>>>

%The naive approach for nested counting in the paper \cite{Denghang2023} is to unwind all inner counting operators. This approach may generate a large number of states, which is linear to the product of the counting bounds in regex and exponential to the length of the regex. For example, naive unwinding regex $(a^{\{1,1000\}})^{\{1,2\}}$ generate a CEFA containing 1000 states. But if we only unwind the outer counting operator of the regex, not the inner, we obtain a CEFA containing only 2 states. The side effect is that we need two registers to store the information of counting in the inner regex (unwinding of the outer counting operator repeats the inner regex, which also repeats the registers of the inner regex). We present a heuristic to reduce the number of states generated by the unwinding process. The idea is to unwind the counting operator with less bound and fewer side effects firstly. We present the heuristic in the following algorithm.

@@

For a regex $\regex$, we can approximate the number of states $size(\regex)$ generated by the following formula: 
\begin{itemize}
  \item If $\regex = \emptyset$ or $\regex = \epsilon$, then $size(\regex) = 0$. 
  \item If $\regex = \regex_1 \cdot \regex_2$ or $\regex = \regex_1 + \regex_2$, then $size(\regex) = size(\regex_1) + size(\regex_2)$.
  \item If $\regex = \regex_1^*$, then $size(\regex) = size(\regex_1)$.
  \item If $\regex = \regex_1 \cap \regex_2$, then $size(\regex) = size(\regex_1)*size(\regex_2)$.
  \item If $\regex = \overline{\regex_1}$, then $size(\regex) = 2^{size(\regex_1)} + 1$.
  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
  \item For $\regex = \regex_1^{\{n,m\}}$ or $\regex = \regex_1^{\{m,\infty\}}$, $size(\regex)$ is equal to $m*size(\regex_1)$ if it is unwinded, and $size(\regex_1)$ otherwise.
\end{itemize}

@@

We unwind the counting operator with less bound first. The reason is that the number of states generated by the counting operator is linear to its bound. If we unwind the counting operator with less bound first, the number of states generated by the unwinding process is less. However, the hardness of solving the $\cefadec$ problem is not only relevant to the number of states, but also the number of registers. For a regex $\regex$, we can approximate the number of registers $reg(\regex)$ generated by the following formula:
\begin{itemize}
  \item If $\regex = \emptyset$, $\regex = \epsilon$, $\regex = \regex_1^*$ or $\regex = \overline{\regex_1}$, then $reg(\regex) = 0$.
  \item If $\regex = \regex_1\cdot \regex_2$ or $\regex = \regex_1 \cap \regex_2$, then $reg(\regex) = reg(\regex_1) + reg(\regex_2)$.
  \item If $\regex = \regex_1 + \regex_2$, then $reg(\regex) = reg(\regex_1) + reg(\regex_2) + 1$.
  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the number of registers can be computed by the formula of $\cap$ and $\bar{}$.
  \item For $\regex = \regex_1^{\{n,m\}}$ or $\regex = \regex^{\{m, \infty\}}$, $reg(\regex)$ is equal to $m*reg(\regex_1)$ if it is unwinded, and $1$ otherwise.
\end{itemize}

@@

\medskip
\noindent
\emph{Step 3. (Unwind the counting operator with less bound and fewer registers first).}

In the above two steps, we compute the number of states and registers generated by each counting operator. In some cases, we can unwind the counting operator with less bound and fewer registers directly. But in most cases, there are multiple counting operators where one counting operator has less bound and another counting operator has fewer registers. To consider the number of states and registers at the same time, we can compute the score of each counting operator by the following formula:
\begin{itemize}
  \item $score(\regex) = size(\regex) * (reg(\regex)+1)$.
\end{itemize} 
The reason why we use the product to combine the two factors is that the hardness of the $\cefadec$ problem is highly relevant to the number of variables in the final existential LIA formula, which is linear to the product of the number of states and the number of registers. Using $reg(\regex) + 1$ not $reg(\regex)$ is because the number of registers may be $0$.We unwind the counting operator with the smallest score first.

\begin{example}
 Consider the regex $(a^{\{1,1000\}})^{\{1,2\}}$. The score of $^{\{1,1000\}}$ is $1000*(0+1) = 1000$ and the score of $^{\{1,2\}}$ is $2*(2+1) = 6$. We unwind the outer counting operator first and obtain a CEFA containing only two states.  
\end{example}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As mentioned in Section~\ref{sec:algorithm}, if some counting operators are in the scope of a complement operator, it is necessary to unfold these counting operators before the complement operation. The exponential blow-up in the unfolding combined with the blow-up in the complement operation will produce a huge number of states. To alleviate this problem, we propose two approximations of the complement operation, i.e., an under-approximation and an over-approximation, and combine them into a procedure that achieves a nice balance between efficiency and precision.

@@

\paragraph*{Under-approximation}
%
One way of doing the complement operation is to transform $e_1$ into a regular expression that does not contain counting operators. We compute an under-approximation of $e$ by replacing each occurrence of the counting operators in $e_1$, say $(e_2)^{\{m,n\}}$ or $(e_2^{\{n, \infty\}})$, by $e_2^*$. Let us use $e'_1$ to denote the resulting regular expression. Note that $e'_1$ does not contain counting operators. Moreover, it is easy to observe that $\Lang(e_1) \subseteq \Lang(e'_1)$. As a result, we have $\Lang(\overline{e'_1}) \subseteq \Lang(\overline{e_1})$. That is, $\overline{e'_1}$ is an under-approximation of $\overline{e_1}$.

%The under-approximation is replacing each counting operator in the subregex of complement with a Kleene star operator. The language of the subregex is a superset of the language of it after under-approximation. In sequential, the complement of the subregex is a subset of the origin. Note that this approach only works for non-nested complement, which means there is no other complement operators in the complement. For nested complement,  we naively handle the inner complement operators as paper \cite{Denghang2023} and then under-approximate the outer complement operator.

@@

Without loss of generality, we can assume that there are no transitions out of $F$ in $\aut_{e_1}$. If this is not the case, then a new state $q_f$ can be introduced to transform $\aut_{e_1}$ into a CEFA satisfying this constraint, where $q_f$ is the unique accepting state. 
Then we define $\cB= (R \cup \{r'\}, Q \cup \{q_s\}, \Sigma, \delta', I, Q \cup \{q_s\}, \alpha')$, where
\begin{itemize}
  \item $r'$ is a new register not in $R$, 
  \item $q_s$ is a new state not in $Q$,
  \item $\delta'$ comprises the following transitions, 
  \begin{itemize}
    \item $(q, a, q', (\vec{v}, 0))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \not \in F$,
%
    \item $(q, a, q_s,(\vec{0}, 0))$ such that there do \emph{not} exist $a$-labeled transitions out of $q$, that is, for every $q'$ and $\vec{v}$, we have $(q, a, q', \vec{v}) \not\in \delta$,
%
    \item $(q, a, q', (\vec{v}, 1))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \in F$,
%
    \item $(q_s, a, q_s, (\vec{0}, 0))$ such that $a \in \Sigma$, 
  \end{itemize}
  \item $\alpha'$ is defined as $r' = 1 \rightarrow \neg \alpha$.
\end{itemize}
Intuitively, $q_s$ is a sink state that deals with the situations where there are no runs, and $r'$ is a register newly introduced so that the different cases of the non-accepting runs of $\aut_{e_1}$ can be captured by a single accepting condition $\alpha'$.

@@

The under-approximation is replacing each counting operator in the subregex of complement with a Kleene star operator. The language of the subregex is a superset of the language of it after under-approximation. In sequential, the complement of the subregex is a subset of the origin. Note that this approach only works for non-nested complement, which means there is no other complement operators in the complement. For nested complement,  we naively handle the inner complement operators as paper \cite{Denghang2023} and then under-approximate the outer complement operator.

@@

$r$ is the register used to check whether the accepted state is also accepted in the original CEFA. $s$ is the self-loop state accepting all strings that are not accepted by the original CEFA. $f$ is the merged accepted state of the original CEFA. If there is a transition $(q, a, q', \vec{v})$ terminating at the accepted state $q'$ in the original CEFA, we add a corresponding transition $(q, a, f, (\vec{v}, 1))$ in the over-approximated complement. From the definition of $\delta'$, we can see that only the transitions goto $f$ increase the value of register $r$. The value of $r$ is $1$ if and only if the accepted state is also accepted in the original CEFA. With the above construction, the meaning of the accepting condition $\alpha'$ is obvious: If the string terminates at the accepting state in the original CEFA, then the accepting condition $\alpha$ is false.

@@

 In the definition of CEFA in section \ref{sec:automaton}, a run $q_0 \xrightarrow[\myvec{\overline{v}_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{\overline{v}_n}]{a_n} q_n$ is \emph{accepting} by $\overline{\aut}_{over}$ if $q_n \in (Q\cup\{s,f\})\backslash F$ and $\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ is true, where $\myvec{\overline{v}'} = \sum \limits_{j \in [n]} \myvec{\overline{v}_j}$. If a string is not accepted by $\overline{\aut}_{over}$, then each run of the string is not accepting run, which means the run terminates at state $q\in F$ or the accepting condition $\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ is $false$.

@@

 If each run makes the accepting condition $\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ false, then the negation of the accepting condition $\neg\alpha'(\myvec{\overline{v}'}/(R\cup\{r\}))$ is true. The negation can be rewriten to $(r=1\wedge \alpha)(\myvec{\overline{v}'}/(R\cup\{r\})) = (r=1)(\myvec{\overline{v}'}/(R\cup\{r\}))\wedge \alpha(\myvec{\overline{v}'}/(R\cup\{r\}))$. In our construction, only the transitions terminate at $f$ increase one to the value of $r$, so that the run must end at $f$ to make $r=1$ be true. $\alpha(\myvec{\overline{v}'}/(R\cup\{r\})) = \alpha(\myvec{v'}/R)$ is also need to be true, implying that the run is an accepting run of $\aut$.
\end{proof}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Recall that in Section~\ref{sec:algorithm}, the satisfiability of the original RECL constraint is reduced to the satisfiability of the existential LIA formula
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
Let us use $\theta$ to denote this formula. 
Then we utilize some SMT solver, say Z3, to solve the satisfiability of $\theta$. If Z3 reports ``sat'', then it also returns a model, namely, an assignment $\amodel$ of natural numbers to the free integer variables in $\theta$. Then we utilize $\amodel$ to generate a model $\amodel'$ of the original RECL constraint $\varphi$, i.e., an assignment of strings to the string variables so that $\varphi$ is evaluated to true under $\amodel'$.

@@

It turns out that this model generation process is inefficient, mainly due to the fact that the information about the occurrences of vectors $\myvec{v}$ in $\cB_i$ in the searching process is too coarse. In the sequel, we would like to refine this process by using the LIA formulas to obtain the information about the number of occurrences of transitions, instead of just the occurrences of integer vectors, and utilize this information to guide the search process. 
Specifically, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, we introduce an integer variable $\anivarz_{i, (q, a, q', \myvec{v})}$ to denote the number of occurrences of the transition, and solve the satisfiability of the following formula,  
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right]\ \wedge \\
\ \ \bigwedge \limits_{i \in [n]} \bigwedge \limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} = \sum \limits_{(q, a, q', \myvec{v}) \mbox{ occurs in } \cB_i} \anivarz_{i, (q, a, q', \myvec{v})}.
\end{array}
\]
If an SMT solver reports that the aforementioned formula is satisfiable, then the solver can also return a model, which assigns to each variable $\anivarz_{i, (q, a, q', \myvec{v})}$ a natural number. 
%For each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, let us use $nt_{i, (q, a, q', \myvec{v})}$ to denote the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model.
Then we utilize these values to generate a model (an assignment of strings to string variables) of $\varphi$. 
Let $i \in [n]$. We use a function $\lambda_i$ to record the number of occurrences of the transitions in $\cB_i$. During the search process for $\cB_i$, each time when a transition $(q, a, q', \myvec{v})$ is traversed, $\lambda_i$ is updated by decrementing $\lambda_i((q, a, q', \myvec{v}))$, moreover, $a$ is also recorded. Initially, for each $i \in [n]$ and each transition $(q, a, q', \myvec{v})$ in $\cB_i$, $\lambda_i((q, a, q', \myvec{v}))$ is set to be the value of $\anivarz_{i, (q, a, q', \myvec{v})}$ in the model. When a final state of $\cB_i$ is reached and $\lambda_i((q, a, q', \myvec{v}))$ becomes zero for each transition $(q, a, q', \myvec{v})$ in $\cB_i$, then a desired string is generated for the string variable $x_i$.

Moreover, propose a method to prune the search space further. Let $i \in [n]$. Suppose that during the search process for $\cB_i$, let $q$ be the current state of $\cB_i$ and $\lambda_i$ record the remaining occurrences of transitions of $\cB_i$. We compute the set of transitions of $\cB_i$ that are reachable from the state $q$. Let $\cT_q$ denote this set. If $\{(q, a, q', \myvec{c}) \mid \lambda_i((q, a, q', \myvec{c})) > 0\} \not \subseteq \cT_q$, then we know that it is impossible to continue the search process so that a string accepted by $\cB_i$ will be found and meanwhile the constraint on the number of transitions enforced by $\lambda_i$ is satisfied. Therefore, in this case, the search process for $\cB_i$ in this direction halts, and the search process is backtracked.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
In the section \ref{sec:algorithm}, we present an algorithm to solve the $\cefadec$ problem. The algorithm computes existential LIA formulas and then uses an SMT solver to solve them. However, this procedure only answers the question of whether the language of the regex is empty or not. If the language is not empty, we need to find a word in the language in practice. In our previous paper \cite{Denghang2023}, we do not explicitly present the algorithm to find a word in the language because of space limitation. We used a naive approach, which randomly searched the CEFAs by the depth-first search based on the register values obtained from solving the existential LIA formula. However, this approach is not efficient when the regexes are non-register-representable. Suppose that there are $n$ registers whose values are $r_1,\cdots, r_n$ respectively in the CEFA. We may visit each transition $O(r_1*r_2\cdots *r_n)$ times during the search. To relieve this problem, we present a heuristic to find a word in the language as follows.

@@

For each CEFA $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$, we already know the values of its registers after solving the existential LIA formula. The values can be seen as a constant vector of length $|R|$. We use $\myvec{R}$ to denote the vector and use $\eta((q, a, q', \myvec{v}))$ to denote the parikh image of $\aut$. Then the transition times visited by the accepted string are obtained from the following formula:
$$ \sum \limits_{(q, a, q', \myvec{v})\in \delta} \eta((q, a, q', \myvec{v}))\myvec{v} = \myvec{R}$$

\medskip
\noindent
\emph{Step 2. (Prune the search tree early by the transition times during the search process).}

After computing the transition times visited by the accepted string, we prune the search tree of the CEFA based on the transition times. The idea is that we only visit the state which is able to reach all transitions need to be visited. Other states are deleted from the CEFA so that we can reduce the search space.
Note that the transition times are updated during the search process, so the search space is reduced dynamically. The pseudo-code is presented in Algorithm~\ref{alg:findword}.
